---
title: "p8105_hw2_zw2541"
author: "Zixu_Wang"
date: "9/30/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1

## Read and clean the data
```{r}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
transit_data
```

This dataset contains such variables: "line", "station name", "station latitude", "stantion longitude", "route1" to "route11", "entry", "vending", "entrance type" and "ADA compliance".

My data cleaning step:

*   Import the dataset
*   Clean the names
*   Select the variables I'm interested in 
*   Use the ifelse function to convert the entry variable from character to a logical variable

This dataset contains `r nrow(transit_data)` rows and `r ncol(transit_data)` columns. And these data are not tidy.