---
title: "p8105_hw2_zw2541"
author: "Zixu_Wang"
date: "9/30/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

## Read and clean the data
```{r}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
transit_data
```

This dataset contains such variables: "line", "station name", "station latitude", "stantion longitude", "route1" to "route11", "entry", "vending", "entrance type" and "ADA compliance".

My data cleaning step:

*   Import the dataset
*   Clean the names
*   Select the variables I'm interested in 
*   Use the ifelse function to convert the entry variable from character to a logical variable

This dataset contains `r nrow(transit_data)` rows and `r ncol(transit_data)` columns. And these data are not tidy.

## Question 1
How many distinct stations are there? 
```{r}
distinct_stations = distinct(transit_data, station_name, line, .keep_all = T) 
nrow(distinct_stations)
```
There are 465 distinct stations.

## Question 2
How many stations are ADA compliant?
```{r}
distinct_stations %>% 
  filter(ada == T) %>% 
  nrow()
```
84 stations are ADA compliant.

## Question 3
What proportion of station entrances / exits without vending allow entrance?

If we only count the distinct stations, the propotion of station entrances / exits without vending allow entrance is
```{r}
distinct_stations %>% 
  filter(vending == "NO", entry == "TRUE") %>% 
  nrow() / nrow(filter(transit_data, vending == "NO"))
```

If we count all of the stations (different entrances of the same unique station as difference stations), the propotion of station entrances / exits without vending allow entrance is
```{r}
transit_data %>% 
  filter(vending == "NO", entry == "TRUE") %>% 
  nrow() / nrow(filter(transit_data, vending == "NO"))
```

## Reformat data
```{r}
reform_transit_data = transit_data %>% 
  gather(key = route_number, value = route_name, route1:route11) 
  
reform_distinct_stations = distinct_stations %>% 
  gather(key = route_number, value = route_name, route1:route11)
```

*  How many distinct stations serve the A train?
```{r}
reform_distinct_stations %>% 
  filter(route_name == "A") %>% 
  nrow()
```
Therefore, 60 distinct stations serve the A train.

*  Of the stations that serve the A train, how many are ADA compliant?
```{r}
reform_distinct_stations %>% 
  filter(route_name == "A", ada == "TRUE") %>% 
  nrow()
```
Therefore, 17 of the stations that serve the A train are ADA compliant.




