---
title: "p8105_hw2_zw2541"
author: "Zixu_Wang"
date: "9/30/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

## Read and clean the data
```{r, message = FALSE}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
transit_data
```

This dataset contains such variables: "line", "station name", "station latitude", "stantion longitude", "route1" to "route11", "entry", "vending", "entrance type" and "ADA compliance".

My data cleaning step:

* Import the dataset
* Clean the names
* Select the variables I'm interested in 
* Use the ifelse function to convert the entry variable from character to a logical variable

This dataset contains `r nrow(transit_data)` rows and `r ncol(transit_data)` columns. And these data are not tidy.

## Questions

* How many distinct stations are there? 
```{r}
distinct_stations = distinct(transit_data, station_name, line, .keep_all = T) 
nrow(distinct_stations)
```
There are 465 distinct stations.

* How many stations are ADA compliant?
```{r}
distinct_stations %>% 
  filter(ada == T) %>% 
  nrow()
```
84 stations are ADA compliant.

* What proportion of station entrances / exits without vending allow entrance?

If we only count the distinct stations, the propotion of station entrances / exits without vending allow entrance is
```{r}
distinct_stations %>% 
  filter(vending == "NO", entry == "TRUE") %>% 
  nrow() / nrow(filter(distinct_stations, vending == "NO"))
```

If we count all of the stations (different entrances of the same unique station as difference stations), the propotion of station entrances / exits without vending allow entrance is
```{r}
transit_data %>% 
  filter(vending == "NO", entry == "TRUE") %>% 
  nrow() / nrow(filter(transit_data, vending == "NO"))
```

## Reformat data
```{r}
reform_transit_data = transit_data %>% 
  gather(key = route_number, value = route_name, route1:route11) 
reform_transit_data
  
reform_distinct_stations = distinct_stations %>% 
  gather(key = route_number, value = route_name, route1:route11)
reform_distinct_stations
```

* How many distinct stations serve the A train?
```{r}
reform_distinct_stations %>% 
  filter(route_name == "A") %>% 
  nrow()
```
Therefore, 60 distinct stations serve the A train.

* Of the stations that serve the A train, how many are ADA compliant?
```{r}
reform_distinct_stations %>% 
  filter(route_name == "A", ada == "TRUE") %>% 
  nrow()
```
Therefore, 17 of the stations that serve the A train are ADA compliant.

# Problem 2

## Read and clean the data

* Read and clean the Mr. Trash Wheel sheet.
```{r}
wheel_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1, range = "A2:N338") %>% 
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = as.integer(round(sports_balls, 0)))
wheel_data
```

* Read and clean precipitation data for 2016.
```{r}
precip_2016 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 5, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = "2016")
```

* Read and clean precipitation data for 2017
```{r}
precip_2017 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = "2017")
```

* Combine datasets and convert month to a character variable
```{r}
precip_data = bind_rows(precip_2016, precip_2017) %>% 
  mutate(month = month.name[month]) %>% 
  select(year, everything())
precip_data
```

## Describe the datasets

The dataset "wheel_data" contains the data from the first sheet of the "HealthyHarborWaterWheelTotals2018-7-28.xlsx" file. This data has `r nrow(wheel_data)` rows, `r ncol(wheel_data)` columns and `r nrow(wheel_data)` observations. Besides, this dataset has `r ncol(wheel_data)` variables. Taking "weight_tons" as an example, dumpster1 collected 4.31 tons of trash on May 16th, 2014.

The dataset "precip_data" combines the data from the forth and fifth sheet of the "HealthyHarborWaterWheelTotals2018-7-28.xlsx" file. This data has `r nrow(precip_data)` rows, `r ncol(precip_data)` colums and `r nrow(precip_data)` observations. Besides, this dataset has `r ncol(precip_data)` variables. For example, the total precipetation in January 2016 is 3.23.

## Questions

* What was the total precipitation in 2017?
```{r}
sum(precip_2017$total)
```
The total precipitation in 2017 is 32.93.

* What was the median number of sports balls in a dumpster in 2016?
```{r}
median(filter(wheel_data, year == "2016")$sports_balls)
```
In 2016, the median number of sports balls in a dumspter is 26.

# Problem3

## Read and clean the data
```{r}
library(p8105.datasets)

data("brfss_smart2010")

brfss_smart2010_tidy = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(excellent_verygood = excellent + very_good)
brfss_smart2010_tidy
```

## Questions

* How many unique locations are included in the dataset? Is every state represented? What state is observed the most?
```{r}
length(unique(brfss_smart2010_tidy$locationdesc))
length(unique(brfss_smart2010_tidy$locationabbr))
brfss_smart2010_tidy %>% 
  count(locationabbr) %>% 
  arrange(desc(n))
```
There are 404 unique locations included in the dataset. Every state is represented and NJ is observed the most.

*In 2002, what is the median of the “Excellent” response value?
```{r}
median(filter(brfss_smart2010_tidy, year == "2002")$excellent, na.rm = TRUE)
```
In 2002, the median of the "Excellent" response value is 23.6

* Make a histogram of “Excellent” response values in the year 2002.
```{r}
brfss_smart2010_tidy %>% 
  filter(year == "2002") %>% 
  ggplot(aes(x = excellent)) + 
  geom_histogram()
```

* Make a scatterplot showing the proportion of "Excellent" response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.
```{r}
brfss_smart2010_tidy %>% 
  filter(locationdesc %in% c("NY - New York County", "NY - Queens County")) %>% 
  ggplot(aes(x = year, y = excellent, color = locationdesc)) + 
  geom_point(size = 5) +
  labs(x = "Year", y = "the Proportion of Excellent Response Values (%)")
```

